\RequirePackage[hyphens]{url}

\documentclass[sigconf]{acmart}
\input{format/i523}
\usepackage{url}
\begin{document}
\title{Big Data Analysis for Wild File Prevention, and Tracking}
\author{Zachary Meier}
\affiliation{%
  \institution{Indiana University}
  \city{Bloomington} 
  \state{Indiana} 
  \postcode{47408}
}
\email{zrmeier@indiana.edu}
% The default list of authors is too long for headers}
\renewcommand{\shortauthors}{Z. Meier et al.}
\begin{abstract}
In Southern California, wildfires never seem too out of place and seem to be a commonly accepted as a norm for the most part.
Over the years this problem has only continued to 
get worse. These fires started to cause massive amounts of property 
damage, loss of life, and hazardous conditions for everyone.  When 
wildfires break out they soon become unpredictable due to wind 
conditions and vegetation.  To amplify problems further, most places 
where burns take place are hard to reach by the emergency crews 
fighting them.  The combination of these factors results in scattered 
resources, playing catch up rather than preventing the spread and 
evacuating large areas.  Using Predictive analytics to not only know 
where a fire is going but, possibly preventing it in the first place 
may help save lives and billions of dollars.
\end{abstract}
\keywords{hid346, i523}
\maketitle
\section{Introduction}
Wildfires are a large part of life in Southern California, and just 
like any other type of event has a large effect on the population, we 
as humans do our best to prepare for it.   Originally, it was building 
houses with stucco and tile roofs so they were less likely to catch 
fire. Now with the advancement of big data, there is a new way to look 
at and defend against wildfires.  These new technology and methods will 
allow first responders to be ahead of the fire.  Allowing for them to 
set backfires, and protect properties more efficiently, as well as help 
prevent new fires popping up from Santa Ana winds which help carry fire 
embers to new locations. There are a couple new technologies that help 
deal with this exact type of problem.
\section{The Data}
Data, the gold of the information era we currently live in. However, in 
some ways it is a much more complex item to mine for.  It is 
ubiquitous, which end up adding more layers to the complexity of 
squiring the gold we are after. To fine the information we are after we 
need to understand what we are looking for.  Most times it is something 
already in place, and needs no new real creation to squire it.  Only by 
manipulating this data, can we get the answers we need.  Below will 
give a brief insight into the data and framework usage.  Then we can 
start to look at how programs are being built to harness this data in 
new, creative, and useful ways. 
\subsection{Capturing Fire}
For years fire has always been seen as something natural and 
unpredictable.  The same we would view a wild animal.  There is a 
certain mystery to it yet it is a force that demands respect.  Though 
the respect of this powerful force of nature may never dwindle.  The 
mysteries of how it moves, burns and it's overall behavior are becoming 
less so.
To know where fire is going to go, and to simulate it correctly you 
need two types of data.  There is a need for static data, and a need 
for dynamic data. These two types are dependent.  If static data were 
to be used only, we would know what a fire would like to burn most 
likely, but that is about it.  However if you layer dynamic data into 
that situation, you create many possibilities that mirror real life 
conditions, and allow for fire fighters to, know where the fire is 
going to spread.
\subsection{Static Data} 
The static data that will most likely be seen in this instance are 
maps.  The understanding of vegetation in the geographic area allows of 
an understanding of what is most likely to fuel the fires spread.  In 
addition there are topographic maps which show elevation of the land, 
which can also help understand how a fire might move through different 
terrains.
\subsection{Dynamic Data}
This is the information captured that can help understand what affects 
fire behavior and spread in a certain amount of time. This data for 
wild fires can be broken up in to two different categories, Graphical 
User Interface (GIS) fire perimeters, and meteorological variables.
\subsubsection{User Interface Fire Perimeters}
In order to predict the fire and it behavior there is a need of 
accurately identifying were it is located and where it is going via 
satellite.  The use of satellites equipped with  MODIS (Moderate 
Resolution Imaging Spec-troradiometer)  Which are used to create high 
resolution pictures of the area while being able to actually see where 
the fire is located without interference from weather conditions and 
smoke \cite{BRUN}.
\subsubsection{Meteorological Variables}
These variables are really the main cause of fire spread.  Many 
environmental factors such as temperature and humidity play a decent 
role in the behavior of the fire.  However, the largest contributing 
factor to fire spread and over all behavior is wind.  When wind speeds 
reduce typically fires are much easier to control. In harnessing data 
about wind conditions in Southern California, there maybe hope for 
creating real time, accurate fire spread data.  Thus helping fire 
fighters react appropriately to where the fire is going, rather than 
just reacting to the the fire itself. therefore, helping save money in 
property damage and live, and resources.

\section{Firemap}
Fire map is an interactive web based application in which you can 
simulate wildfire models.  It also utilizes WIFIRE's data sets.  It 
allows users to explore the map and work with multiple layers.  The 
layers can be changed, or have their opacities changed, allowing for 
multiple layers to be built upon one another.  This give the user 
customizable data and better insight in the data. 
\subsection{Pylaski}
To make Firemap to be interactive and useful to users it must be able 
to provide all the different types data from the multiple sources where 
the data is located.  To do this Pylaski uses a rest service to query 
internal as well as external sources for date.  It send a query for 
each individual request and then returns all the data at once back to 
the client.  Pylaski is also able to use different schema's as well as 
different formats.  This works fairly well creating an idea prediction.  
However as it is collecting massive amounts of data it is unable to 
query all of the data request because of size limitations.  Though 
limiting it is most likely able to satisfy the requirements of the user 
\cite{CRAWLFiremap}. 
\subsection{GeoServer}
GeoServer is a open source server for geographical datasets.  It allows 
you to get data from a large number of geospatial data collection 
agencies.  With these you can pull up a large number of maps.
\begin{description}
\item [Historical Fires:] These are file that were collected on 
previous fire parameters.  There were recoded for the whole United 
States from the years 2000-2015.  In addition there are many records 
from California specifically from 1878-2014.  All which if clicked on, 
tell you descriptions of the fire in detail.
\item [Satellite Detection:] Used thermal imaging and MODIS satellites 
to produce high resolution photos of fires from above.  These fly over 
about two time a day, and are uploaded to NASA.
\item [Smoke Areas:] Show smoke plumes and help possibly identify 
the start of new wildfires. Data comes from NOAA
\item [Red Flag Areas:] Red flag areas are places where conditions are 
right for a possible fire occur. These are typically indicated by high 
wind speed or low relative humidity.  this data comes from the 
National Wildfire Coordinating Group.
\item [Census Block:]The census block comes from the US Census which 
keep track of population density.  Allowing for consideration of people 
affected by a fire.
\item [Surface Fuels:] Used thirteen ways to categorize fuel for fires 
through vegetation.  This is a large factor in how fires will spread as 
they need easy to burn fuel to keep burning.
\item [Canopy Cover:] The percentage of forest floor covered by trees.  
Crutal in analyzing model crown fires which spread from top layers of 
trees.
\item [Camera Viewsheds:] Geographical areas with visibility of surrounding areas in which the HPWREN and AlertTahoe cameras are used to visualize the area of the fire.
\end{description}
\section{WIFIRE}
Wildfires in the right setting are beneficial to health of an 
ecosystem.  However, when you put millions of people within that 
ecosystem, it becomes a problem.  This is the problem that the WIFIRE 
project started at University of California San Diego is trying to 
remedy.   The have created an end to end cyber infrastructure system to 
create real time prediction, visualization for wildfires in San Diego 
Counter. They use real time data from sensors and satellite for 
observation as well as meteorological study in order to create up to 
date predictions.  Being that San Diego county sees many wildfires, it 
is a prime location for testing this project out.  The most recent fire 
in 2007 cost over one billion dollars and causes half a million people 
to be evacuated from their homes\citep{NPR}. In order to collect the 
correct and most up to date information for tracking a wild fire they 
call upon the High Performance Wireless Research and Education Network 
known as HPWREN.  In addition, the use various partner sites that also 
collect large amounts of data.  These come from various types of data, 
such as, meteorology, vision and audio and hydrology.   To make all of 
relevant and precise for the end users, such as the fire fighters.
\subsection{Systems}
Their systems for collecting this heterogeneous data uses a few 
different layers of architecture. There is a pipeline composed of many 
different sensors from the San Diego county, such as ground based 
sensors, SDG\& E (San Diego Gas \& Electric), Water Meteorology 
Stations and the HPWREN as well as collect satellite imagery.  With all 
this heterogeneous data coming in the needed a way to process all the 
data.  They achieved this with  an open source workflow system called 
Kepler which organized the execution of the real time data processed.  
Then that system distributed it to a portal for end user consumption.  
This system is meant to handle large data input as well as execute 
computation in parallel so all data given is up to date 
\cite{ALTINTAS}.
\subsection{Subsystems}
The most important part of all this system is the data pipeline.  Which 
in their workflow is called the Data Communication Subsystem.  This 
system handles the large input of data and works with integration many 
different data sets and information from imaging, whether it be 
satellite or camera.  The influx of data is come is produced by three 
types of communication systems.  Meteorology stations, still cameras 
around San Diego County, and Satellite and Aerial data.  This data 
communication layer uses REST services to take in the data to the 
database and the who process is done in memory to keep processing time 
as low as possible.
\subsection{Data Mining}
Big data today would not be what it is today if not for the fundamental 
process of data mining.  Predicting trends based out of huge amounts of 
data, in order to understand exploit that data for practical use and 
application. WIFIRE also uses this methodology to find data relating to 
Santa Ana winds.  Analyzing the data from the meteorological stations 
from around the county can help them realize when and where a Santa Ana 
wind possible. This data would be useful to the firefighter, because it 
would allow them to know the fire may have a chance to spread 
further.  In addition this data can also be put into fire modeling to 
help predict wildfires movement even better in the future.
\subsection{Visualization}
Using visual this project's goal is provide the most transparent data 
to first responders. Using satellite imagery to understand areas that 
have been previously scarred by fires and less likely to burn, while 
identifying areas that may be more susceptible to burning.  This uses 
the topography and analization from previous fires to give them a 
better picture of how fires moved throughout the county.
In addition to using the aerial data, the project is trying to create 
an application for pedestrians and bystanders to use to help track 
active fires from the ground.  The premise of this idea is that if 
enough users take photos they can recreate where that photo was taken 
in a 3D space and track the fire that way.  It would also help identify 
the fires over all perimeter.  Once again helping first responders, 
more effectively fight the fires and know the whole area it covers and 
where it is heading.
\subsection{Assimilation}
Most programs on the market are based off of just pure data, and 
produce a simulation base of that.  The problem with that is fires and 
still not predictable.  Using the behavior of real fires from real time 
data, and compares that against stimulation's, to correct the errors 
based off of the simulation alone. The system uses recursive algorithms 
created by Extended Kalman Filter and a Ensemble Kalman Filter.   It 
uses existing fire models such as those in FARSITE in a separate 
workflow.  Allowing for the real time fire date to run in process, and 
compare them against the simulation data previously provided.  This 
allows them to be compared and move them to the update step, which is 
where the data assimilation takes place.  Correcting previous errors 
for the simulation, allowing the simulations for future predictions to 
be much more accurate.
\subsection{Santa Ana Winds}
Due to the amount of wildfires that spur up per year in the area, the 
amount of data received over wildfire may very.   However, one of the 
biggest factors in wildfires and their spreading ability is with Santa 
Ana Winds.  They are based off a couple key conditions, such as 
direction, speed and  relative humidity.  All of which create extremely 
dangerous situations for fires.  Luckily there are many sensor around 
San Diego County that can detect this happening.  The issues is 
however, is when an alert is sent, it is only for a specific point.  
With the kepler workflow they are able to take in data from those 
individual stations and run it through a program called WindNinja which 
can take vegetation, topography and weather stations into account 
produce a mapping of the areas affected.  These are ran through the 
workflow breaking down tiles from the stations in 50km by 50km blocks 
for marking the wind occurrence.  This then runs through a data 
processing software such as Spark or Hadoop.

 
\section{Conclusion}
In conclusion, the use of big data in fire prevention and tracking is 
an impressive system.  It incorporates a multitude of sensors and data
from various sources.  Creating predictions of fires and creating data 
models which will help first responders saves lives, and protect 
property from damage.  This application will allow efficient of 
firefighter resources and allow for them to be one step ahead of the 
fire.  Which until now was impossible.  Only when the weather changed, 
were we able to change and react.   With the predictive modeling and 
machine learning, the models will continue to get more accurate as the 
topography and fire interactions continue into the future.  While this 
system will not prevent all wild fires, it will be instrumental in 
slowing them down and containing them.  Once again man will try to tame fire as we have tried to do since the beginning of our existence. 
\begin{acks}
The authors would like to thank Dr. Gregor von Laszewski and the TA's for their help and support.
\end{acks}
\bibliographystyle{ACM-Reference-Format}
\bibliography{report}
\end{document}
